---
title: "Buying the Farm: The Distribution of Local and Cultural Benefits from Farmland Conservation"
author: "Amy Ando, Frederick Nyanzu, and Bryan Parthum"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
    code_folding:  hide
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, echo=FALSE}
library(readxl)
library(mlogit)
library(gmnl)
library(data.table)
library(here)
library(tidyverse)
library(magrittr)
library(kableExtra)
```

# Summary

This document provides an overview of a simulation of the designed discrete choice experiment exploring local and cultural values of farmland conservation. The experiment design was creasted using [-dcreate-](https://ideas.repec.org/c/boc/bocode/s458059.html) (Hole, 2015) in [Stata 15](https://www.stata.com/stata15/). Models are estimated in [Rstudio](https://rstudio.com/), an open source platform using the [-gmnl-](https://www.jstatsoft.org/article/view/v079i02) package (Sarrias and Daziano, 2017). 

# The Experiment Design

The experiment design consists of 3 blocks of 8 choice cards (occasions)---a total of 26 possible cards. Each choice occasion has a status-quo (no policy) option and an alternative (policy) option (i.e. dichotomous choice). 

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
## Read experiment design from generated design
data <- read_excel('../experiment_design/store/design_matrix.xlsx')

## Summary table of attribute levels
design_matrix_miles <- data %>%
                       filter(treatment=="miles") %>%
                       mutate(block = ifelse(cost==0,'Status Quo',block),
                              miles = paste(dist,'','miles')) %>%
                       group_by(block) %>%
                       distinct(cost,nature,farmland,food,meals,fruit,veggies,miles) %>%
                       arrange(factor(block, levels = c('Status Quo',1,2,3)))

design_matrix_minutes <- data %>%
                         filter(treatment=="minutes") %>%
                         mutate(block = ifelse(cost==0,'Status Quo',block),
                                minutes = paste(dist,'','minutes')) %>%    
                         group_by(block) %>%
                         distinct(cost,nature,farmland,food,meals,fruit,veggies,minutes) %>%
                         arrange(factor(block, levels = c('Status Quo',1,2,3))) %>%
                         ungroup()

design_matrix <- cbind(design_matrix_miles, (design_matrix_minutes %<>% select(minutes))) %>% relocate(minutes, .after = miles) %>% relocate(food, .before = meals) 

kable(design_matrix, caption='Unique alternatives in each of the 3 blocks',booktabs = TRUE, align = c('l','c','c','c')) %>% kable_classic() %>% kable_styling(font_size = 16) %>%
      footnote('There are two treatments (minutes and miles) that have identical designs',footnote_as_chunk = T)
```

# Simulation Data

To explore if our experiment design provides the necessary variation to estimate the model, we simulate a completed survey using random choices of 600 individuals. Because choices are random, the coefficients will (should) be insignificant. 

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

## Repeat for arbitrary number of respondents, assume one respondent per block-treatment pair 
## Three blocks of eight cards with two alternatives and two treatments in initial design
data <- bind_rows(replicate(100, data, simplify = FALSE))

## Number of individuals in sample
nrow(data)/12

## Generate unique individual identifiers
data %<>% mutate(ind_id = ceiling(row_number()/12))

## This includes both treatments. Create a unique card identifier for a pooled (both treatments) dataset
data %<>% group_by(treatment,block,card) %>%
          mutate(pooled_card_id = cur_group_id())

## Generate unique individual-card identifier (unique choice occasions)
data %<>% group_by(ind_id,pooled_card_id) %>%
          mutate(ind_card_id = cur_group_id())

## Generate random choices
set.seed(42)
data %<>% group_by(ind_card_id) %>%
          mutate(choice = sample(c(0,1), replace=FALSE, size=2)) 

## Generate alternative specific constant
data %<>% mutate(asc = ifelse(alt==2,1,0))

## Generate interactions
data %<>% mutate(food_fruit = food * fruit)

## Write to CSV 
write_csv(data,'../store/simulation_data.csv')
```

# Preference-space Models

We first estimate the model using a simple multinomial logit (MNL). We then introduce the random parameter model (MXL) with both uncorrelated random parameters and fully correlated. We assume the distribution on _cost_ is fixed in order to derive subsequent empirical distributions of willingness to pay (WTP) using the delta method. This means that the marginal utility of money is uniform across respondents.

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

## Create mlogit data
d <-  mlogit.data(data,
                  # shape='long', 
                  drop.index = TRUE,
                  id.var='ind_id', ## unique to individual_id
                  chid.var = 'ind_card_id', ## unique to individual_id and card_id
                  choice='choice',
                  shape='long',
                  alt.var='alt') ## the number alternative on each card
                  # opposite=c('bid')) ## if changing distribution of cost coefficient to log-normal
```

## Null Model

$$V_{it} = ACS + \varepsilon_{it}$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

####################################################
#######################################  Null Models
####################################################

null_model <- mlogit(choice ~ asc | 0 , data=d)
summary(null_model)
```

## Multinomial Logit

$$V_{it} = ACS + \lambda cost_{it} + \beta_1 nature_{it} + \beta_2 farmland_{it} + \beta_3 food_{it} + \beta_4 distance_{it} + \varepsilon_{it}$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

####################################################
################################################  MNL
####################################################

pref_clogit <-  gmnl(choice ~ asc + cost + nature + farmland + food + dist | 0 ,
                     data=d,
                     model='mnl')

summary(pref_clogit)
```

## Random Parameter Logit

$$V_{it} = ACS + \lambda_i cost_{it} + \beta_{i1} nature + \beta_{i2} farmland + \beta_{i3} food + \beta_{i4} distance + \varepsilon_{it}$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

pref_uncorr <-  gmnl(choice ~ asc + cost + nature + farmland + food + dist | 0 ,
                    data=d,
                    ranp=c(cost='n',nature='n',farmland='n',food='n',dist='n'),
                    model='mixl',
                    panel=TRUE,
                    correlation=FALSE,
                    seed=42)

summary(pref_uncorr)
```
