---
title: "Buying the Farm: The Distribution of Local and Cultural Benefits from Farmland Conservation"
author: "Amy Ando, Corey Lang, Frederick Nyanzu, and Bryan Parthum"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
    code_folding:  hide
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, echo=FALSE}
library(readxl)
library(mlogit)
library(gmnl)
library(data.table)
library(here)
library(tidyverse)
library(magrittr)
library(kableExtra)
```

# Summary

This document provides an overview of a simulation of the designed discrete choice experiment exploring local and cultural values of farmland conservation. The experiment design was created using [-dcreate-](https://ideas.repec.org/c/boc/bocode/s458059.html) (Hole, 2015) in [Stata 15](https://www.stata.com/stata15/). Models are estimated in [Rstudio](https://rstudio.com/), an open-source platform using the [-gmnl-](https://www.jstatsoft.org/article/view/v079i02) package (Sarrias and Daziano, 2017). 

# The Experiment Design

The experiment design consists of 7 blocks of 6 choice cards (occasions)---a total of 42 possible cards. Each choice occasion has a status-quo (no policy) option and an alternative (policy) option (i.e. dichotomous choice). 

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
## Read experiment design from generated design
data <- read_excel('..\\experiment_design\\store\\design_matrix.xlsx')
## Summary table of attribute levels
design_matrix_miles <- data %>%
                       filter(treatment=="miles") %>%
                       mutate(block = ifelse(cost==0,'Status Quo',block),
                              miles = paste(distance,'','miles')) %>%
                       group_by(block) %>%
                       distinct(cost,nature,farmland,meals_nature,meals_farmland,miles) %>%
                       arrange(factor(block, levels = c('Status Quo',1,2,3,4,5,6))) %>% 
                       mutate(card = seq(n())) %>%
                       mutate(card = ifelse(block=='Status Quo',0,card)) %>% 
                       ungroup()
design_matrix_minutes <- data %>%
                         filter(treatment=="minutes") %>%
                         mutate(block = ifelse(cost==0,'Status Quo',block),
                                minutes = paste(distance,'','minutes')) %>%    
                         group_by(block) %>%
                       distinct(cost,nature,farmland,meals_nature,meals_farmland,minutes) %>%
                         arrange(factor(block, levels = c('Status Quo',1,2,3,4,5,6))) %>% 
                       mutate(card = seq(n())) %>%
                       mutate(card = ifelse(block=='Status Quo',0,card)) %>% 
                       ungroup()
design_matrix <- cbind(design_matrix_miles, (design_matrix_minutes %<>% select(minutes))) %>% relocate(minutes, .after = miles) %>% relocate(card, .after = block)
kable(design_matrix, caption='Unique alternatives in each of the 6 blocks',booktabs = TRUE, align = c('l','c','c','c')) %>% kable_classic() %>% kable_styling(font_size = 16) %>%
      footnote('There are two treatments (minutes and miles) that have identical designs',footnote_as_chunk = T)
```

# Simulation Data

To explore if our experiment design provides the necessary variation to estimate the model, we simulate a completed survey using random choices of 350 individuals. Because choices are random, the coefficients will (should) be insignificant. The miles treatment was selected for this simulation.

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
## Repeat for arbitrary number of respondents, assume one respondent per block-treatment pair 
## Three blocks of eight cards with two alternatives and two treatments in initial design
data %<>% filter(treatment=='miles')
data <- bind_rows(replicate(50, data, simplify = FALSE))
## Number of individuals in sample
nrow(data)/12
## Generate unique individual identifiers
data %<>% mutate(ind_id = ceiling(row_number()/12))
## This includes both treatments. Create a unique card identifier for a pooled (both treatments) dataset
data %<>% group_by(treatment,block,card) %>%
          mutate(pooled_card_id = cur_group_id())
## Generate unique individual-card identifier (unique choice occasions)
data %<>% group_by(ind_id,pooled_card_id) %>%
          mutate(ind_card_id = cur_group_id())
## Generate random choices
set.seed(42)
data %<>% group_by(ind_card_id) %>%
          mutate(choice = sample(c(0,1), replace=FALSE, size=2)) 
## Generate alternative specific constant
data %<>% mutate(asc = ifelse(alt==2,1,0))
## Generate interactions
# data %<>% mutate(meals_nature = ifelse(nature>0,meals_nature/nature,0),
#                  meals_farmland = ifelse(farmland>0,meals_farmland/farmland,0))
## Generate interactions
data %<>% mutate(distance_nature = distance * nature,
                 distance_farmland = distance * farmland,
                 distance_meals_nature = distance * meals_nature,
                 distance_meals_farmland = distance * meals_farmland)
## Write to CSV 
write_csv(data,'..\\store\\simulation_data.csv')
```

# Preference-space Models

We first estimate the model using a simple multinomial logit (MNL). Then we explore additional heterogeneity in this model by adding interactions between $distance$ and $nature$, $farmland$, $meals^{nature}$, and $meals^{farm}$, and then nonlinear preferences across the attribute space. We then introduce the random parameter model (MXL) with uncorrelated random parameters. We then extend the random parameter model to a model with interactions and then, lastly, we relax the assumption that random parameters are uncorrelated to recover a full covariance matrix. We assume the distribution on $cost$ is fixed in order to derive subsequent empirical distributions of willingness to pay (WTP) using the a Krinsky-Robb bootstrap method. This means that the marginal utility of money is uniform across respondents.

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
## Create mlogit data
d <-  mlogit.data(data,
                  # shape='long', 
                  drop.index = TRUE,
                  id.var='ind_id', ## unique to individual_id
                  chid.var = 'ind_card_id', ## unique to individual_id and card_id
                  choice='choice',
                  shape='long',
                  alt.var='alt') ## the number alternative on each card
                  # opposite=c('bid')) ## if changing distribution of cost coefficient to log-normal
```

## Null Model

$$V_{it} = ASC + \varepsilon_{it}$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
####################################################
#######################################  Null Models
####################################################
null_model <- mlogit(choice ~ asc | 0, data=d)
summary(null_model)
```

## Multinomial Logit

$$V_{it} = ASC + \lambda cost_{it} + \beta^1 nature_{it} + \beta^2 farmland_{it} + \beta^3 meals^{nature}_{it} + \beta^4 meals^{farm}_{it} + \beta^5 distance_{it} + \varepsilon_{it}$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
####################################################
################################################  MNL
####################################################
pref_clogit <-  gmnl(choice ~ asc + cost + nature + farmland + meals_nature + meals_farmland + distance | 0 ,
                     data=d,
                     model='mnl')
summary(pref_clogit)
```

## Multinomial Logit with Interactions

In this simulation, we expand the standard multinomial model (with preference parameters fixed across the population) to include interactions with distance. This allows preferences to vary in the distance to each attribute in the model. A random parameter model was not estimable in these simulations.

$$
\begin{aligned}
V_{it} = & \ ASC + \lambda cost_{it} + \beta^1 nature_{it} + \beta^2 farmland_{it} + \beta^3 meals^{nature}_{it} + \beta^4 meals^{farm}_{it} + \beta^5 distance_{it} \\ 
& + \beta^6(distance_{it} \times nature_{it}) + \beta^7(distance_{it} \times farmland_{it}) \\
& + \beta^8(distance_{it} \times meals^{nature}_{it}) + \beta^9(distance_{it} \times meals^{farm}_{it}) + \varepsilon_{it}
\end{aligned}
$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
pref_interact <-  gmnl(choice ~ asc + cost + nature + farmland + meals_nature + meals_farmland + distance + distance_nature + distance_farmland + distance_meals_nature + distance_meals_farmland | 0 ,
                    data=d,
                    model='mnl')
summary(pref_interact)
```

## Nonparametric Multinomial Logit

Here, the parameters on cost ($\lambda$) and distance ($\beta^5$) ar assumed to be linear (implying a constant marginal utility of income and distance) while the other parameters correspond to each attribute's level $n$. 

<!-- During this simulation, the only nonlinear parameters that could be estimated were $\beta^1_n$ and $\beta^2_n$, the parameters associated with $nature$ and $farmland$. Attempting to estimate the others (either independantly or jointly) resulted in a computationally singular matrix (not estimable). Perhaps this has to do with the constraints imposed on the design? In particular, the $meals$ attributes are constrained to be positive only when the $acres$ attributes are also positive. This is a strong condition to impose on the design. Prior to imposing this condition, all $\beta^i_n$ parameters were estimable. Perhaps it would be worth thinking about which parameters are most relevant to estimate nonlinearly. $Acres$? $Distance$? It is also entirely possible that this could be remedied with real choice data and a reasonable sample size. -->

$$V_{it} = ASC + \lambda cost_{it} + \sum_n \beta^1_{n} nature_{it} + \sum_n \beta^2_{n} farmland_{it} + \sum_n \beta_n^3 meals^{nature}_{it} + \sum_n \beta_n^4 meals^{farm}_{it} +  \beta^5 distance_{it} + \varepsilon_{it}$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
pref_factor <-  gmnl(choice ~ asc + cost + as.factor(nature) + as.factor(farmland) + as.factor(meals_nature) + as.factor(meals_farmland) + distance | 0 ,
                    data=d,
                    model='mnl')
summary(pref_factor)
```

## Random Parameter Logit

In this simulation, we relax the assumption that parameters are distributed equally across the population and allow for random preferences. The parameter on cost ($\lambda$) remains fixed, while others are now denoted with the subscript $i$ representing individual $i$. Parameters are assumed uncorrelated and distributed normal.

$$V_{it} = ASC + \lambda cost_{it} + \beta^1_{i} nature_{it} + \beta^2_{i} farmland_{it} + \beta^3_{i} meals^{nature}_{it} + \beta^4_{i} meals^{farm}_{it} + \beta^5_{i} distance_{it} + \varepsilon_{it}$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

pref_uncorr <-  gmnl(choice ~ asc + cost + nature + farmland + meals_nature + meals_farmland + distance + distance_nature + distance_farmland + distance_meals_nature + distance_meals_farmland | 0 ,
                    data=d,
                    ranp=c(nature='n',farmland='n',meals_nature='n',meals_farmland='n',distance='n',distance_nature='n',distance_farmland='n',distance_meals_nature='n',distance_meals_farmland='n'),
                    model='mixl',
                    panel=TRUE,
                    correlation=FALSE,
                    seed=42)

summary(pref_uncorr)
```

## Random Parameter Logit with Interactions

Here, we extend the previous random parameter model to incorporate interactions between $distance$ and other attributes.

$$
\begin{aligned}
V_{it} = & \ ASC + \lambda cost_{it} + \beta_i^1 nature_{it} + \beta_i^2 farmland_{it} + \beta_i^3 meals^{nature}_{it} + \beta_i^4 meals^{farm}_{it} + \beta_i^5 distance_{it} \\ 
& + \beta_i^6(distance_{it} \times nature_{it}) + \beta_i^7(distance_{it} \times farmland_{it}) \\
& + \beta_i^8(distance_{it} \times meals^{nature}_{it}) + \beta_i^9(distance_{it} \times meals^{farm}_{it}) + \varepsilon_{it}
\end{aligned}
$$
```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

pref_uncorr_int <-  gmnl(choice ~ asc + cost + nature + farmland + meals_nature + meals_farmland + distance | 0 ,
                    data=d,
                    ranp=c(nature='n',farmland='n',meals_nature='n',meals_farmland='n',distance='n'),
                    model='mixl',
                    panel=TRUE,
                    correlation=FALSE,
                    seed=42)

summary(pref_uncorr_int)
```

## Random Parameter Logit with Interactions and Correlated Random Parameters

This is our final specification in preference-space. Random parameters are now assumed to be fully-correlated.

$$V_{it} = ASC + \lambda cost_{it} + \beta^1_{i} nature_{it} + \beta^2_{i} farmland_{it} + \beta^3_{i} meals^{nature}_{it} + \beta^4_{i} meals^{farm}_{it} + \beta^5_{i} distance_{it} + \varepsilon_{it}$$

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

pref_corr <-  gmnl(choice ~ asc + cost + nature + farmland + meals_nature + meals_farmland + distance + distance_nature + distance_farmland + distance_meals_nature + distance_meals_farmland | 0 ,
                    data=d,
                    ranp=c(nature='n',farmland='n',meals_nature='n',meals_farmland='n',distance='n',distance_nature='n',distance_farmland='n',distance_meals_nature='n',distance_meals_farmland='n'),
                    model='mixl',
                    panel=TRUE,
                    correlation=TRUE,
                    seed=42)

summary(pref_corr)
```